Optimal tensor denoising and completion based on ordinal observations

### 2020-01-28 ICML/miaoyan/ordinalT.tex
update draft
update

\begin{algorithm}[tb]
        \caption{Ordinal tensor decomposition }\label{alg}
        \begin{algorithmic}[]
            \STATE{\bfseries Input:}  \text{ Ordinal tensor }
            $\mathcal{Y}\in [L]^{d_1\times\cdots\times d_K}
            $,\\ \hspace{.43in} Rank $\mr\in \mathbb{N_+}^{K}$,\\
            \hspace{.43in} Entry-wise bound $\alpha\in \mathbb{R_+}$.
            \STATE{\bfseries Output:} $(\hat\Theta,\hat{\mb}) =  \argmax_{(\Theta,\mb)\in \mathcal{D}\times\mathcal{B}}  \mathcal{L}_{\mathcal{Y},\Omega}(\Theta,\mb).$
            \STATE Initialize  Core tensor $\mathcal{C}^{(0)}$,\\
            \hspace{.5in} Factor matrices $ \{\mM_1^{(0)},\cdots,\mM_K^{(0)}\}$,\\
            \hspace{.5in} Cut-off points $\mb^{(0)}$.
            \FOR{$t = 1,2,\cdots,$}
            \FOR{$k = 1,2,\cdots,K$}
            \STATE{Update $\mM_k$ while fixing other blocks.}
            \STATE $\mM_k^{(t+1)}\gets\argmax_{\mM_k\in\mathbb{R}^{d_k\times r_K}}\mathcal{L}_{\mathcal{Y},\Omega}(\mM_k)$
            \newline
             s.t. $\mnormSize{}{\Theta^{(t+1)}}\leq \alpha$, where $\Theta^{(t+1)}$ is the parameter tensor based on the current block estimates. 
            \ENDFOR
            \STATE {Update $\mathcal{C}$ while fixing other blocks: $\mathcal{C}^{(t+1)} \gets \argmax_{\mathcal{C}\in \mathbb{R}^{r_1\times\cdots\times r_K}}\mathcal{L}_{\mathcal{Y},\Omega}(\mathcal{C})$, s.t. $\mnormSize{}{\Theta^{(t+1)}}\leq \alpha$. 
               \STATE {Update $\Theta$ based on the current block estimates: $\Theta^{(t+1)} \gets \mathcal{C}^{(t+1)}\times_1\mM_1^{(t+1)}\cdots\times_K\mM_K^{(t+1)}$. 
            \STATE {Update $\mb$ while fixing $\Theta^{(t+1)}$: $\mb^{(t+1)} \gets \argmax_{\mb\in \mathbb{R}^{L-1}}\mathcal{L}_{\mathcal{Y},\Omega}\big(\Theta^{(t+1)},\mb\big)$
            \ENDFOR
            \STATE \textbf{return}
            $\Theta,\mb$
        \end{algorithmic}
    \end{algorithm}

@@

\section{Experiments}
In this section, we evaluate the empirical performance of our rank-constrained M-estimatation. We consider ordinal tensors with various number of categories, and compare the recovery accuracy with other tensor-based methods. Unless otherwise stated, we generate ordinal-valued tensor under the latent-variable model~\eqref{eq:latent} with i.i.d.\ $N(0,\sigma^2)$ noise. The signal tensor is simulated based on~\eqref{eq:Tucker}, with core tensor generated from i.i.d. N(0,1), and the factor matrices uniformly from orthogonal ensembles. The cut-off points are generated i.i.d.from $N(0,1)$. In each simulation study, we report the summary statistics across $n_{\text{sim}}$ = 30 replications.

@@

\begin{table}[H]
\resizebox{\linewidth}{!}{%
\begin{tabular}{c|ccc|c}
\hline
\multirow{2}{*}{Critera}&\multicolumn{3}{c|}{Our method}&Continuous method\\
& $\hat y^{(\text{Mean})}_\omega$ & $\hat y^{(\text{Median})}_\omega$ & $\hat y^{(\text{Mode})}_\omega$ &~\cite{filipovic2015tucker}\\
\hline
MSE& &  & 0.1504&0.1604\\
MAD&  &  &0.1504&0.1600\\
MCR& &&0.1502&0.1598\\
\hline
\end{tabular}
}
\caption{Comparison of tensor denoising on HCP dataset.}
\end{table}

>>>>

\begin{algorithm}[tb]
        \caption{Ordinal tensor decomposition }\label{alg}
        \begin{algorithmic}[]
            \STATE{\bfseries Input:}  \text{ Ordinal tensor }
            $\mathcal{Y}\in [L]^{d_1\times\cdots\times d_K}
            $,\\ \hspace{.43in} Rank $\mr\in \mathbb{N}_+^{K}$,\\
            \hspace{.43in} Entry-wise bound $\alpha\in \mathbb{R_+}$.
            \STATE{\bfseries Output:} $(\hat\Theta,\hat{\mb}) =  \argmax_{(\Theta,\mb)\in \mathcal{D}\times\mathcal{B}}  \mathcal{L}_{\mathcal{Y},\Omega}(\Theta,\mb).$
            \STATE Initialize  Core tensor $\mathcal{C}^{(0)}$,\\
            \hspace{.5in} Factor matrices $ \{\mM_1^{(0)},\cdots,\mM_K^{(0)}\}$,\\
            \hspace{.5in} Cut-off points $\mb^{(0)}$.
            \FOR{$t = 1,2,\cdots,$}
            \FOR{$k = 1,2,\cdots,K$}
            \STATE{Update $\mM_k$ while fixing other blocks.}
            \STATE $\mM_k^{(t+1)}\gets\argmax_{\mM_k\in\mathbb{R}^{d_k\times r_K}}\mathcal{L}_{\mathcal{Y},\Omega}(\mM_k)$
            \newline
             s.t.\ $\mnormSize{}{\Theta^{(t+1)}}\leq \alpha$, where $\Theta^{(t+1)}$ is the parameter tensor based on the current block estimates. 
            \ENDFOR
            \STATE {Update $\mathcal{C}$ while fixing other blocks: $\mathcal{C}^{(t+1)} \gets \argmax_{\mathcal{C}\in \mathbb{R}^{r_1\times\cdots\times r_K}}\mathcal{L}_{\mathcal{Y},\Omega}(\mathcal{C})$, s.t.\ $\mnormSize{}{\Theta^{(t+1)}}\leq \alpha$. 
               \STATE {Update $\Theta$ based on the current block estimates: $\Theta^{(t+1)} \gets \mathcal{C}^{(t+1)}\times_1\mM_1^{(t+1)}\cdots\times_K\mM_K^{(t+1)}$. 
            \STATE {Update $\mb$ while fixing $\Theta^{(t+1)}$: $\mb^{(t+1)} \gets \argmax_{\mb\in \mathbb{R}^{L-1}}\mathcal{L}_{\mathcal{Y},\Omega}\big(\Theta^{(t+1)},\mb\big)$
            \ENDFOR
            \STATE \textbf{return}
            $\Theta,\mb$
        \end{algorithmic}
    \end{algorithm}

@@

\section{Experiments}
In this section, we evaluate the empirical performance of our rank-constrained estimation. We consider ordinal tensors with various number of categories, and compare the recovery accuracy with other tensor-based methods. Unless otherwise stated, we generate ordinal-valued tensor under the latent-variable model~\eqref{eq:latent} with i.i.d.\ $N(0,\sigma^2)$ noise. The signal tensor is simulated based on~\eqref{eq:Tucker}, with core tensor generated from i.i.d. N(0,1), and the factor matrices uniformly from orthogonal ensembles. The cut-off points are generated i.i.d.from $N(0,1)$. In each simulation study, we report the summary statistics across $n_{\text{sim}}$ = 30 replications.

@@

%\begin{table}[H]
%\resizebox{\linewidth}{!}{%
%\begin{tabular}{c|ccc|c}
%\hline
%\multirow{2}{*}{Critera}&\multicolumn{3}{c|}{Our method}&Continuous method\\
%& $\hat y^{(\text{Mean})}_\omega$ & $\hat y^{(\text{Median})}_\omega$ & $\hat y^{(\text{Mode})}_\omega$ &~\cite{filipovic2015tucker}\\
%\hline
%MSE& &  & 0.1504&0.1604\\
%MAD&  &  &0.1504&0.1600\\
%MCR& &&0.1502&0.1598\\
%\hline
%\end{tabular}
%}
%\caption{Comparison of tensor denoising on HCP dataset.}
%\end{table}
